---
title: "Project"
output: pdf_document
---

```{r, message=FALSE}
library(flexsurv)
library(coxed)
library(survival)
library(simsurv)
library(MASS)
library(tidyverse)
library(nph)
```
## Proportional and non-proportional models under different right-sensoring and their plots

## Proportional hazard model
### High right-censoring, maxt = 5
```{r}
set.seed(123456)
N = 1000 ## number of patients/observations
covs = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

s1 = simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), maxt = 5) ## Simulate survival data

dat = merge(covs, s1)
mod <- flexsurvspline(Surv(eventtime, status) ~ trt, data = dat) ## Create survival object

plot(survfit(Surv(eventtime, status) ~ trt, data = dat), main = "Proportional model", ylab = "Survival probability",
     xlab = "Time") 

logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 0 )
logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 1 )
logrank.test(dat$eventtime, dat$status, dat$trt, rho = 1, gamma = 0 )
logrank.maxtest(dat$eventtime, dat$status, dat$trt)
```

Log-rank test: the Chi-Squared test statistic is 32.7 with 1 degree of freedom and the corresponding p-value is 1.1e-8, we reject the null hypothesis.

Log-rank test for a late effect: the Chi-Squared test statistic is 283 with 1 degree of freedom and the corresponding p-value is 6.1e-8, we reject the null hypothesis.

Log-rank test for an early effect: the Chi-Squared test statistic is 62 with 1 degree of freedom and the corresponding p-value is 9.4e-10, we reject the null hypothesis.

All three log-rank tests successfully reject the null hypothesis. The typr I errors are the p-values of each test. The type II error is 0.136.


### Lower right-censoring: maxt = 7.5
```{r}
set.seed(123456)
N = 1000 ## number of patients/observations
covs = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

s1 = simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), maxt = 7.5) ## Simulate survival data

dat = merge(covs, s1)
mod <- flexsurvspline(Surv(eventtime, status) ~ trt, data = dat) ## Create survival object

plot(survfit(Surv(eventtime, status) ~ trt, data = dat), main = "Proportional model", ylab = "Survival probability",
     xlab = "Time") 

logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 0 )
logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 1 )
logrank.test(dat$eventtime, dat$status, dat$trt, rho = 1, gamma = 0 )
logrank.maxtest(dat$eventtime, dat$status, dat$trt)
```

Log-rank test: the Chi-Squared test statistic is 52,1 with 1 degree of freedom and the corresponding p-value is 5.3e-13, we reject the null hypothesis.

Log-rank test for a late effect: the Chi-Squared test statistic is 258 with 1 degree of freedom and the corresponding p-value is 2.0e-12, we reject the null hypothesis.

Log-rank test for an early effect: the Chi-Squared test statistic is 122 with 1 degree of freedom and the corresponding p-value is 5.2e-10, we reject the null hypothesis.

All three log-rank tests successfully reject the null hypothesis. The typr I errors are the p-values of each test. The type II error is 0.270.


### Even lower right-censoring: maxt = 10
```{r}
set.seed(123456)
N = 1000 ## number of patients/observations
covs = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

s1 = simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), maxt = 10) ## Simulate survival data

dat = merge(covs, s1)
mod <- flexsurvspline(Surv(eventtime, status) ~ trt, data = dat) ## Create survival object

plot(survfit(Surv(eventtime, status) ~ trt, data = dat), main = "Proportional model", ylab = "Survival probability",
     xlab = "Time") 

logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 0 )
logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 1 )
logrank.test(dat$eventtime, dat$status, dat$trt, rho = 1, gamma = 0 )
logrank.maxtest(dat$eventtime, dat$status, dat$trt)
```

Log-rank test: the Chi-Squared test statistic is 52.4 with 1 degree of freedom and the corresponding p-value is 4.5e-13, we reject the null hypothesis.

Log-rank test for a late effect: the Chi-Squared test statistic is 202 with 1 degree of freedom and the corresponding p-value is 4.3e-11, we reject the null hypothesis.

Log-rank test for an early effect: the Chi-Squared test statistic is 137 with 1 degree of freedom and the corresponding p-value is 3.3e-10, we reject the null hypothesis.

All three log-rank tests successfully reject the null hypothesis. The typr I errors are the p-values of each test. The type II error is 0.365.



## Non-proportional hazard model
### High right-censoring, maxt = 5
```{r}
set.seed(123456)
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

pars_tde = c(trt = 0.4)

s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), tde = pars_tde, tdefunction = 'log', maxt = 5) 

dat2 =  merge(s2, covs2)
mod2 <- flexsurvspline(Surv(eventtime, status) ~ trt, data = dat2)

plot(survfit(Surv(eventtime, status) ~ trt, data = dat2), main = "Non-proportional model", ylab = "Survival probability",
     xlab = "Time") 

logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1 )
logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0 )
logrank.maxtest(dat2$eventtime, dat2$status, dat2$trt)
```

Log-rank test: the Chi-Squared test statistic is 2.87 with 1 degree of freedom and the corresponding p-value is 0.09, so we fail to reject the null hypothesis.

Log-rank test for a late effect: the Chi-Squared test statistic is 19.1 with 1 degree of freedom and the corresponding p-value is 0.98, so we fail to reject the null hypothesis.

Log-rank test for an early effect: the Chi-Squared test statistic is 5.99 with 1 degree of freedom and the corresponding p-value is 0.01, so we reject the null hypothesis.

Therefore, when the hazard model is time-dependent, only the log-rank test for an early effect successfully rejects the null hypothesis. The typr I errors are the p-values of each test. The type II error is 0.538.


### Lower right-censoring, maxt = 7.5
```{r}
set.seed(123456)
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

pars_tde = c(trt = 0.4)

s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), tde = pars_tde, tdefunction = 'log', maxt = 7.5) 

dat2 =  merge(s2, covs2)
mod2 <- flexsurvspline(Surv(eventtime, status) ~ trt, data = dat2)

plot(survfit(Surv(eventtime, status) ~ trt, data = dat2), main = "Non-proportional model", ylab = "Survival probability",
     xlab = "Time") 

logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1 )
logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0 )
logrank.maxtest(dat2$eventtime, dat2$status, dat2$trt)
```

Log-rank test: the Chi-Squared test statistic is 0.805 with 1 degree of freedom and the corresponding p-value is 0.37, so we fail to reject the null hypothesis.

Log-rank test for a late effect: the Chi-Squared test statistic is 3.03 with 1 degree of freedom and the corresponding p-value is 0.35, so we fail to reject the null hypothesis.

Log-rank test for an early effect: the Chi-Squared test statistic is 2.16 with 1 degree of freedom and the corresponding p-value is 0.02, so we reject the null hypothesis.

Therefore, when the hazard model is time-dependent, only the log-rank test for an early effect successfully rejects the null hypothesis. The typr I errors are the p-values of each test. The type II error is 0.838.


### Even lower right-censoring, maxt = 10
```{r}
set.seed(123456)
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

pars_tde = c(trt = 0.4)

s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), tde = pars_tde, tdefunction = 'log', maxt = 10) 

dat2 =  merge(s2, covs2)
mod2 <- flexsurvspline(Surv(eventtime, status) ~ trt, data = dat2)

plot(survfit(Surv(eventtime, status) ~ trt, data = dat2), main = "Non-proportional model", ylab = "Survival probability",
     xlab = "Time") 

logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1 )
logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0 )
logrank.maxtest(dat2$eventtime, dat2$status, dat2$trt)
```

Log-rank test: the Chi-Squared test statistic is 0.0755 with 1 degree of freedom and the corresponding p-value is 0.79, so we fail to reject the null hypothesis.

Log-rank test for a late effect: the Chi-Squared test statistic is 0.241 with 1 degree of freedom and the corresponding p-value is 0.07, so we fail to reject the null hypothesis.

Log-rank test for an early effect: the Chi-Squared test statistic is 0.22 with 1 degree of freedom and the corresponding p-value is 0.03, so we reject the null hypothesis.

Therefore, when the hazard model is time-dependent, only the log-rank test for an early effect successfully rejects the null hypothesis. The typr I errors are the p-values of each test. The type II error is 0.908.




## Repeat simulation 50 times under minimized censoring to find type 1&2 error

### Type 2 error
```{r}
## Type 2 error under non-proportional model with an early effect
index = 0
index_early = 0
index_late = 0
n = 50
for (i in 1:n) {
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

pars_tde = c(trt = 0.4)

s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs2, betas = c(trt = -0.5), tde = pars_tde, tdefunction = 'log', maxt = 10) 

dat2 =  merge(s2, covs2)
Test = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
Test_early = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0)
Test_late = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1)

test_df = Test$test
test_early_df = Test_early$test
test_late_df = Test_late$test

p_val = test_df$p
p_val_early = test_early_df$p
p_val_late = test_late_df$p

if (p_val < 0.05) {
index <- index + 1
  }
if (p_val_early < 0.05) {
index_early <- index_early + 1
  }
if (p_val_late < 0.05) {
index_late <- index_late + 1
  }
}

1 - index/n 
1 - index_early/n
1 - index_late/n
```
When the hazard model simulates an early effect, the type 2 error of the unweighted log-rank test is 0.9, the type 2 error of the log-rank test for an early effect is 0.42, and the type 2 error of the log-rank test for a late effect is 0.58. 


```{r}
## Type 2 error under non-proportional model with a late effect
index = 0
index_early = 0
index_late = 0
n = 50
for (i in 1:n) {
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

pars_tde = c(trt = 0.6)

s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs2, betas = c(trt = -0.5), tde = pars_tde, tdefunction = 'log', maxt = 10) 

dat2 =  merge(s2, covs2)

Test = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
Test_early = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0)
Test_late = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1)

test_df = Test$test
test_early_df = Test_early$test
test_late_df = Test_late$test

p_val = test_df$p
p_val_early = test_early_df$p
p_val_late = test_late_df$p

if (p_val < 0.05) {
index <- index + 1
  }
if (p_val_early < 0.05) {
index_early <- index_early + 1
  }
if (p_val_late < 0.05) {
index_late <- index_late + 1
  }
}

1 - index/n 
1 - index_early/n
1 - index_late/n
```
When the hazard model simulates a late effect, the type 2 error of the unweighted log-rank test is 0.18, the type 2 error of the log-rank test for an early effect is 0.98, and the type 2 error of the log-rank test for a late effect is 0.02.

```{r}
## Type 2 error under proportional model
index = 0
index_early = 0
index_late = 0
n = 50
for (i in 1:n){
N = 1000 ## number of patients/observations
covs = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

s1 = simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = -0.5), maxt = 10) ## Simulate survival data

dat = merge(covs, s1)

Test = logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 0 )
Test_early = logrank.test(dat$eventtime, dat$status, dat$trt, rho = 1, gamma = 0)
Test_late = logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 1)

test_df = Test$test
test_early_df = Test_early$test
test_late_df = Test_late$test

p_val = test_df$p
p_val_early = test_early_df$p
p_val_late = test_late_df$p

if (p_val < 0.05) {
index <- index + 1
  }
if (p_val_early < 0.05) {
index_early <- index_early + 1
  }
if (p_val_late < 0.05) {
index_late <- index_late + 1
  }
}

1 - index/n 
1 - index_early/n
1 - index_late/n
```
When the hazard model is proportional, the type 2 error of the unweighted log-rank test is 0, the type 2 error of the log-rank test for an early effect is 0, and the type 2 error of the log-rank test for a late effect is 0.


### Type 1 error
```{r}
## Type 1 error under proportional model
index = 0
index_early = 0
index_late = 0
n = 50
for (i in 1:n){
N = 1000 ## number of patients/observations
covs = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation

s1 = simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = c(trt = 0), maxt = 10) ## Simulate survival data

dat = merge(covs, s1)

Test = logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 0 )
Test_early = logrank.test(dat$eventtime, dat$status, dat$trt, rho = 1, gamma = 0)
Test_late = logrank.test(dat$eventtime, dat$status, dat$trt, rho = 0, gamma = 1)

test_df = Test$test
test_early_df = Test_early$test
test_late_df = Test_late$test

p_val = test_df$p
p_val_early = test_early_df$p
p_val_late = test_late_df$p

if (p_val < 0.05) {
index <- index + 1
  }
if (p_val_early < 0.05) {
index_early <- index_early + 1
  }
if (p_val_late < 0.05) {
index_late <- index_late + 1
  }
}

index/n 
index_early/n
index_late/n
```
When the hazard model is proportional, the type 1 error of the unweighted log-rank test is 0.06, the type 1 error of the log-rank test for an early effect is 0.02, and the type 1 error of the log-rank test for a late effect is 0.02.

```{r}
## Type 1 error under non-proportional model with an early effect
index = 0
index_early = 0
index_late = 0
n = 50
for (i in 1:n) {
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation
pars_tde = c(trt = 0.4)
s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = NULL, tde = pars_tde, tdefunction = 'log', maxt = 10) 
dat2 =  merge(s2, covs2)

Test = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
Test_early = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0)
Test_late = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1)

test_df = Test$test
test_early_df = Test_early$test
test_late_df = Test_late$test

p_val = test_df$p
p_val_early = test_early_df$p
p_val_late = test_late_df$p

if (p_val < 0.05) {
index <- index + 1
  }
if (p_val_early < 0.05) {
index_early <- index_early + 1
  }
if (p_val_late < 0.05) {
index_late <- index_late + 1
  }
}

index/n
index_early/n
index_late/n
```
When the hazard model simulates an early effect, the type 1 error of the unweighted log-rank test is 0.02, the type 1 error of the log-rank test for an early effect is 0.06, and the type 1 error of the log-rank test for a late effect is 0.02.

```{r}
## Type 1 error under non-proportional model with a late effect
index = 0
index_early = 0
index_late = 0
n = 50
for (i in 1:n) {
N = 1000 ## number of patients/observations
covs2 = data.frame(id = 1:N,
                  trt = rbinom(N, 1, 0.5)) ## Covariates of each observation
pars_tde = c(trt = 0.6)
s2 <- simsurv(lambdas = 0.1, gammas = 1.5, x = covs, betas = NULL, tde = pars_tde, tdefunction = 'log', maxt = 10) 
dat2 =  merge(s2, covs2)

Test = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 0 )
Test_early = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 1, gamma = 0)
Test_late = logrank.test(dat2$eventtime, dat2$status, dat2$trt, rho = 0, gamma = 1)

test_df = Test$test
test_early_df = Test_early$test
test_late_df = Test_late$test

p_val = test_df$p
p_val_early = test_early_df$p
p_val_late = test_late_df$p

if (p_val < 0.05) {
index <- index + 1
  }
if (p_val_early < 0.05) {
index_early <- index_early + 1
  }
if (p_val_late < 0.05) {
index_late <- index_late + 1
  }
}

index/n
index_early/n
index_late/n
```
When the hazard model simulates a late effect, the type 1 error of the unweighted log-rank test is 0.04, the type 1 error of the log-rank test for an early effect is 0.04, and the type 2 error of the log-rank test for a late effect is 0.06.


